Lesson 4 (two variables)
========================================================

load the package and the data:
```{r}
library(ggplot2)
pf <- read.csv("~/programme/data_mining/R/Udacity/pseudo_facebook.tsv", sep="\t", header = TRUE, row.names = 1)
summary(pf)
```

scatter plot:
```{r fig.width=7, fig.height=6}
qplot(x=age, y=friend_count, data=pf)
```

similar plot created by ggplot
```{r}
ggplot(aes(x=age, y=friend_count), data=pf) + geom_point()
ggplot(aes(x=age, y=friend_count), data=pf) + geom_point() + xlim(13,90)

ggplot(aes(x=age, y=friend_count), data=pf) + 
  geom_point(alpha = 1/20) + 
  xlim(13,90)

ggplot(aes(x=age, y=friend_count), data=pf) + 
  geom_jitter(alpha = 1/20) + 
  xlim(13,90) 

ggplot(aes(x=age, y=friend_count), data=pf) + 
  geom_point(alpha = 1/20) + 
  xlim(13,90) + 
  coord_trans(y = 'sqrt')

ggplot(aes(x=age, y=friend_count), data=pf) + 
  geom_point(alpha = 1/20, position = position_jitter(h=0)) + 
  xlim(13,90) + 
  coord_trans(y = 'sqrt')
```

friends initiated:
```{r}
ggplot(aes(x=age, y=friendships_initiated), data=pf) + 
  geom_point(alpha = 1/20) + 
  xlim(13,90) + 
  coord_trans(y = 'sqrt')
```

Create table with median using package dplyr:
```{r}
install.packages('dplyr')
library(dplyr)

age_groups <- group_by(pf, age)
pf.fc_by_age <- summarize(age_groups, 
          friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n = n())

pf.fc_by_age <- arrange(pf.fc_by_age, age)
head(pf.fc_by_age)
```

Alternative using the chain of functions:
```{r}
pf.fc_by_age <- pf %.%
  group_by(age) %.%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %.%
  arrange(age)

head(pf.fc_by_age)
```

Plot the dependencies between age and mean friend count
```{r}
qplot(x=age, y=friend_count_mean, data = pf.fc_by_age)
ggplot(aes(x=age,y=friend_count_mean), data=pf.fc_by_age, xlim=c(13,110)) + 
  geom_line() 
```

Overlay of different plots:
```{r}
ggplot(aes(x=age, y=friend_count), data=pf) + 
  xlim(13,90) +
  geom_point(alpha =0.05,
             position=position_jitter(h=0),
             color = 'orange') +
  coord_trans(y='sqrt') +
  geom_line(stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', fun.y = quantile, probs = .1, linetype = 2, color = 'blue') + 
  geom_line(stat = 'summary', fun.y = quantile, probs = .5, linetype = 1, color = 'blue') +
  geom_line(stat = 'summary', fun.y = quantile, probs = .9, linetype = 2, color = 'blue')
```

Lesson zooming in using coord_cartesian:
```{r}
ggplot(aes(x=age, y=friend_count), data=pf) + 
  coord_cartesian(xlim=c(13,90), ylim=c(0,1000)) +
  geom_point(alpha =0.05,
             position=position_jitter(h=0),
             color = 'orange') +
  geom_line(stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', fun.y = quantile, probs = .1, linetype = 2, color = 'blue') + 
  geom_line(stat = 'summary', fun.y = quantile, probs = .5, linetype = 1, color = 'blue') +
  geom_line(stat = 'summary', fun.y = quantile, probs = .9, linetype = 2, color = 'blue')
```

correlation of age and friend_count:
```{r}
cor.test(pf$age, pf$friend_count, method ="pearson")

with(subset(pf, age <= 70), cor.test(age, friend_count))
```

Likes received vs. www_likes received:
```{r}
cor.test(pf$www_likes_received, pf$likes_received, method="pearson")
ggplot(aes(x=www_likes_received, y=likes_received), data=pf) +
    geom_line(stat = 'summary', fun.y = mean, color = "blue") +
    geom_point(alpha =0.5) +
    xlim(0, quantile(pf$www_likes_received, 0.95)) + 
    ylim(0, quantile(pf$likes_received, 0.95)) +
    geom_smooth(method = 'lm', color = "red")

```

