Lesson 5 Reshaping the data
========================================================

loading the data
```{r}
library(ggplot2)
pf <- read.csv("~/programme/data_mining/R/Udacity/pseudo_facebook.tsv", sep="\t", header = TRUE, row.names = 1)
summary(pf)
```

Third option for qualitative variables:
```{r}
library('dplyr')

pf.fc_by_age_gender <- pf %.%
  filter(!is.na(gender)) %.%
  group_by(age, gender) %.%
  summarize(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n = n()) %.%
  ungroup() %.%
  arrange(age)

head(pf.fc_by_age_gender)
```

Plotting third qualitative variable:
```{r}
ggplot(aes(x=age, y=median_friend_count), 
       data=pf.fc_by_age_gender) +
  geom_line(aes(color=gender))
```

Reshaping the data using the dcast function. Result will be a new data frame. The acast function will be used when the result shall be a list or an array. 
```{r}
install.packages('reshape2')
library(reshape2)

pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender,
                                  age ~ gender,
                                  value.var = 'median_friend_count')
```


Plot the ratio of the female to male median friend counts using the new data frame:
```{r}
ggplot(aes(x=age, y=(female / male)), data=pf.fc_by_age_gender.wide) +
  geom_line() + 
  geom_hline(yintercept = 1, alpha = 0.3, linetype = 2)
```

Adding a new variable to the data frame for time sind logged on to fb
```{r}
pf["year_joined"] <- NA
pf$year_joined <- floor((2014*365 - pf$tenure) / 365 )

summary(pf$year_joined)
table(pf$year_joined)
```

cut the data into buckets using the cut function:
```{r}
pf$year_joined.bucket <- cut(pf$year_joined, 
                             c(2004,2009,2011,2012,2014))
head(pf$year_joined.bucket)
table(pf$year_joined.bucket, useNA = 'ifany')
```

