Lesson 4 tests
========================================================

1. scatterplot of price vs x
```{r}
summary(diamonds)
names(diamonds)

cor.test(diamonds$price, diamonds$x, method = 'pearson')
cor.test(diamonds$price, diamonds$y, method = 'pearson')
cor.test(diamonds$price, diamonds$z, method = 'pearson')

ggplot(aes(x=x, y=price), 
       data=diamonds) +
    geom_point(alpha = 0.5)

ggplot(aes(x=price, y=x), 
       data=diamonds) +
    geom_point(alpha = 0.5) + 
    geom_smooth()
```

2. Create a simple scatter plot of price vs depth.
```{r}
cor.test(diamonds$price, diamonds$depth, method="pearson")

ggplot(aes(x=depth, y=price), 
       data=diamonds) +
    geom_point()

ggplot(aes(x=depth, y=price), 
       data=diamonds) +
    geom_point(alpha = 0.01) +
    scale_x_continuous(limits = c(54, 70), breaks=seq(54,70,2))
```

3. Create a scatterplot of price vs carat and omit the top 1% of price and carat
```{r}
diamonds.subset <- subset(diamonds, carat < quantile(carat, 0.99) & price < quantile(price, 0.99))

ggplot(aes(x=price, y=carat), 
       data=diamonds.subset) +
    geom_point(alpha = 0.1) 
```

4. Create a scatterplot of price vs. volume (x * y * z).
Create a new variable for volume in the diamonds data frame.
```{r}
library("gridExtra")

diamonds.extended <- diamonds
diamonds.extended["volume"] <- NA
diamonds.extended$volume <- diamonds.extended$x * diamonds.extended$y *
 diamonds.extended$z

cor.test(diamonds.extended$volume, diamonds.extended$price, method="pearson")

p1 <- ggplot(aes(x=price, y=volume), 
      data=diamonds.extended) +
      geom_point() 

p2 <- ggplot(aes(x=price, y=volume), 
       data=diamonds.extended) +
      scale_y_continuous(limits = c(0, 1000)) +
    geom_point(alpha = 0.5) 

p3 <- ggplot(aes(x=price, y=volume), 
       data=diamonds.extended) +
      scale_y_continuous(limits = c(0, 750)) +
    geom_point(alpha = 0.1) +
    geom_smooth()

grid.arrange(p1, p2, p3, ncol=1)
```

5. Correlation of a subset
Excluding diamonds of 0 volume and with volume higher or equal than 800  
```{r}
diamonds.extended.subset <- subset(diamonds.extended, volume > 0 & volume < 800)

cor.test(diamonds.extended.subset$volume, diamonds.extended.subset$price, method="pearson")

ggplot(aes(x=price, y=volume), 
    data=diamonds.extended.subset) +
    scale_y_continuous(limits = c(0, 750)) +
    geom_point(alpha = 0.1) +
    scale_x_sqrt() + 
    geom_smooth(method="lm")

```

6. Create a new data frame which should contain the following variables in this order.
(1) mean_price
(2) median_price
(3) min_price
(4) max_price
(5) n
```{r}
library(dplyr)

diamond_groups <- group_by(diamonds, clarity)
diamonds.new <- summarize(diamond_groups, 
          mean_price = mean(price),
          median_price = median(price),
          min_price = min(price, na.rm = TRUE),
          max_price = max(price, na.rm = TRUE),
          n = n())
```

7. creating bar plots to verify the data:
```{r}
library(gridExtra)

diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))

p1 <- ggplot(aes(x=clarity, y=mean_price),data=diamonds_mp_by_clarity) +
  geom_bar()

p2 <- ggplot(aes(x=color, y=mean_price),data=diamonds_mp_by_color) +
  geom_bar()

grid.arrange(p1, p2, ncol=1)
```

