---
title: "Sales_Order_Example"
author: "GJA"
date: "11. Juni 2015"
output: html_document
---
## Load the data:
```{r}
SalesOrders1 <- read.csv("~/Downloads/SalesOrders1.csv", sep=";")
summary(SalesOrders1)
```

Important is to ananlyze the structure of the data. For this purpose, the command str() has to be used.
Show structure of the data:
```{r, echo=FALSE}
str(SalesOrders1)
```

Structure shows, that the date is represented as a factor. 
To correct, that date is represented as factor, the stringAsFactor() has to be set to false during upload:
```{r}
SalesOrders1 <- read.csv("~/Downloads/SalesOrders1.csv", sep=";", stringsAsFactors = FALSE)
str(SalesOrders1)
```
Strings are represented as a vector of strings.
Convert string to date:
```{r}
SalesOrders1$DeliveryDate <- as.Date(SalesOrders1$DeliveryDate, format="%Y-%m-%d")
str(SalesOrders1)
```

## Analysis of the raw data:
The analysis of the raw data loaded is the basis for the development of the model. 

1. first step is to analyse the distribution of the different variables in the data set. In order to analyze more than one attribute, the command par() and mfrow has to be used to split the output screen into different sections.
```{r}
par(mfrow=c(2,2))
names(SalesOrders1)
```

Command par() is used to set graphical parameters and mfrow(2,2) splits the output window into four parts. Command names() returns the names of the attributes. 
Historgamms are used to show the distribution of the frequency of the occurrence of an attribute: 
```{r}
hist(SalesOrders1$BuyerRef)
hist(SalesOrders1$DeliveryDate, "weeks")
hist(SalesOrders1$TotalPrice) 
hist(SalesOrders1$PaymentDuration)
```

Histogramms are one dimensional plots.
The histogram for PaymentDuration shows a small peak around 5 and a plateau around 19. This can be seen as indication for a difference in customers payment history. 
Scatter-plots are used to analyze the dependency between two attributes:
```{r}
plot(SalesOrders1$BuyerRef, SalesOrders1$PaymentDuration)
smoothScatter(SalesOrders1$BuyerRef, SalesOrders1$PaymentDuration)
smoothScatter(SalesOrders1$TotalPrice, SalesOrders1$PaymentDuration)
plot(SalesOrders1$TotalPrice, SalesOrders1$PaymentDuration, cex=0.3)
```

The first plot shows the dependency between the attributes PaymentDuration and BuyerRef. The standard function plot() creates for every pair of values one data point. For a large number of data points, the method is not useful, as every single point has to be drawn and the information is for a large number of points not helpful. A possiblity is to reduce the size of the data points in the standard plot with the parameter cex. 
Smooth scatter plots shows the concentration of data points and are more helpful for a large number of records. 

Another possiblity to visualize information about the data within one picture is the Scatterplot-Matrix:
```{r}
library(car)
scatterplotMatrix(SalesOrders1, diagonal="histogram", cex=0.5, smooth=FALSE, reg.line=FALSE)
```

The scatterplot matrix is the central plotting method of the package car (Companion to Applied Regression). 

## Development of the Model
From the visualization of the data no special distribution and dependencies of BuyerRef, DeliveryDate and TotalPrice could be detected. Only the attribut PaymentDuration showed indications, that the payment varies for the customers. 

### Time Series:
First step to create a time series for the payment history per customer ist the creation of a TimeSeries-object:
```{r}
Pd.ts <- aggregate(SalesOrders1$PaymentDuration, by=list(Customer=SalesOrders1$BuyerRef), FUN=ts)
names(Pd.ts) <- c("Customer", "TimeSeries")
str(Pd.ts[1:3,])
```

The function aggregate() processes the data frame record by record aggregates the column PaymentDuration by customer and 
