Udacity Lesson 3
========================================================

Markdown including all commands for lesson 3 of the Exploratory Data Analysis Course at Udacity.

Check the current working directory and list files:
```{r}
getwd()
list.files()
```

Read the data from the file: 
```{r}
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
```

Check the variables by using the names command:
```{r}
names(pf)
```

Histrogram of Users birthdays 1:
```{r}
library(ggplot2)
qplot(x = dob_day, data = pf) 
```

Histrogram of Users birthdays 2 with days of month on the x-axis:
```{r}
library(ggplot2)
qplot(x = dob_day, data = pf) +
  scale_x_discrete(breaks=1:31) 
```

Histrogram of Users birthdays 3:
```{r}
library(ggplot2)
qplot(x = dob_day, data = pf) +
  scale_x_discrete(breaks=1:31) +
  facet_wrap(~dob_month, ncol = 3)
```

Histogram of freind counts:
```{r}
qplot(x = friend_count, data = pf) 
```

Histogram of freind counts with limitation of the x axis:
```{r}
qplot(x = friend_count, data = pf, binwidth = 25, xlim = c(0,1000)) 
```

Histogram of freind counts with limitation of the x axis:
```{r}
qplot(x = friend_count, data = pf, binwidth = 25) +
  scale_x_continuous(limits = c(0,1000), breaks = seq(0,1000,50))
```

Histogram of freind counts with limitation of the x axis with differentiation of gender:
```{r}
qplot(x = friend_count, data = pf, binwidth = 25) +
  scale_x_continuous(limits = c(0,1000), breaks = seq(0,1000,50)) +
  facet_wrap(~gender, ncol = 3)
```

Histogram of freind counts with limitation of the x axis with differentiation of gender omitting NA observations:
```{r}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 25) +
  scale_x_continuous(limits = c(0,1000), breaks = seq(0,1000,50)) +
  facet_wrap(~gender, ncol = 3)
```

Statistics by gender
```{r}
table(pf$gender)
by(pf$friend_count, pf$gender, summary)
```

Tenure by month (30 days):
```{r}
qplot(x = tenure, data = pf, color=I("black"), fill=I('#099DD9'), binwidth = 30)
```

Tenure by year (365 days):
```{r}
qplot(x = tenure/365, data = pf, 
      xlab="Number of years using Facebook",
      ylab="Number of users in sample",
      color=I("black"), fill=I('#F79420'), binwidth = .25) +
  scale_x_continuous(breaks = seq(1,7,1), limits=c(0,7)) 
```

Users by age:
```{r}
qplot(x = age, data = pf, 
      xlab="Age of Facebook users",
      ylab="Number of users in sample",
      color=I("black"), fill=I('#F79420'), binwidth = 1) +
  scale_x_continuous(breaks = seq(0,200,10))
```

Various plots using gridExtra:
```{r}
# load library gridExtra
library(gridExtra)

# define individual plots
plot1 <- qplot(x = friend_count, data = pf, binwidth = 10, xlim = c(0,1000))
plot2 <- qplot(x = log10(friend_count+1), data = pf, binwidth = .05, xlim = c(0,5))
plot3 <- qplot(x = sqrt(friend_count), data = pf, binwidth = .25, xlim = c(0,50))

# arrange plots in grid
grid.arrange(plot1, plot2, plot3, ncol=2)
```

Second variation for log-based x-axis:
```{r}
logScale <- qplot(x = log10(friend_count + 1), data = pf, binwidth = .05)

countScale <- ggplot(aes(x = friend_count), data = pf) +
  geom_histogram() + 
  scale_x_log10()

logScale2 <- qplot(x = friend_count, data = pf) +
  scale_x_log10()

grid.arrange(logScale, countScale, logScale2, ncol = 1)
```

Frequency polygons:
```{r}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth=10) +
  scale_x_continuous(lim=c(0,1000), breaks = seq(0,1000,50)) +
  facet_wrap(~gender)

qplot(x = friend_count, y = ..count../sum(..count..), 
      data = subset(pf, !is.na(gender)), 
      xlab = "Friend Count",
      ylab = "Proporation of Users with that friend count",
      binwidth=10, geom='freqpoly', color = gender) +
  scale_x_continuous(lim=c(0,1000), breaks = seq(0,1000,50)) 


```

Frequency polygon of www_likes:
```{r}
likesScale <- qplot(x = www_likes, 
      data = subset(pf, !is.na(gender)), 
      xlab = "WWW Likes",
      ylab = "Number of users",
      binwidth=100, geom='freqpoly', color = gender) +
  scale_x_continuous(lim=c(20,1500), breaks = seq(20,1500,100)) 

likesLog <- qplot(x = www_likes, 
      data = subset(pf, !is.na(gender)), 
      xlab = "WWW Likes",
      ylab = "Number of users",
      geom='freqpoly', color = gender) +
  scale_x_log10()

grid.arrange(likesScale, likesLog, ncol = 1)
```

Statistics totals of www_likes
```{r}
#table(pf$www_likes)
by(pf$www_likes, pf$gender, sum)
```

Box-plots:
```{r}
option1 <- qplot(x = gender, y = friend_count, 
      data = subset(pf, !is.na(gender)),
      geom='boxplot', color = gender, ylim = c(0,1000)) 

option2 <- qplot(x = gender, y = friend_count, 
      data = subset(pf, !is.na(gender)),
      geom='boxplot', color = gender) + 
  scale_y_continuous(lim=c(0,1000), breaks = seq(0,1000,100)) 

option3 <- qplot(x = gender, y = friend_count, 
      data = subset(pf, !is.na(gender)),
      geom='boxplot', color = gender) + 
  coord_cartesian(ylim=c(0,1000))

grid.arrange(option1, option2, option3, ncol=1)
```

Friend count analysis using boxplot and summary:
```{r}
qplot(x = gender, y = friend_count, 
      data = subset(pf, !is.na(gender)),
      geom='boxplot', color = gender) + 
  coord_cartesian(ylim=c(0,250))

by(pf$friend_count, pf$gender, summary)
```

Friendships initiated analysis using boxplot and summary:
```{r}
qplot(x = gender, y = friendships_initiated, 
      data = subset(pf, !is.na(gender)),
      geom='boxplot', color = gender) + 
  coord_cartesian(ylim=c(0,250))

by(pf$friendships_initiated, pf$gender, summary)
```

Using additional variables:
```{r}
summary(pf$mobile_likes)
summary(pf$mobile_likes > 0)

mobile_check_in <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1, 0)
pf$mobile_check_in <- factor(pf$mobile_check_in)
summary(pf$mobile_check_in)
sum(pf$mobile_check_in == 1) / length(pf$mobile_check_in)
```

